Seniority and Hierarchy Configuration Interaction for Radicals and \\ Excited States

### 2023-10-22 Manuscript/Response_Letter/Response_Letter.tex

{Figures 3 and 4: The data is shown with respect to the number of determinants, which makes the
reader to evaluate the different methods efficiently. However, for readability, I think it will also be
helpful to provide n in eCIn, sCIn, and hCIn in these figures.
}

\alert{
We have experimented including the values of $n$ to the figures, as suggested by the reviewer.
However, we consider that they become somewhat clutered and less readable, and therefore we prefer to leave them unchanged.
With the intention of helping the reader, we have been supplemented the following sentence to the captions of Figures 3 (4):
``Each point denotes one CI model, according to the sequences: HF, hCI1, hCI1.5, hCI2, etc. (green); HF, CIS, CISD, etc. (red); and sCI0, sCI2, and sCI4 (sCI1, sCI3, sCI5) (blue).''
}

{In sections IV-C and IV-D: The EN2 corrections are not discussed at all in the excited states, making
it unnatural compared to the presentation of the radical systems. While it might be challenging to
perform the additional calculations for all the systems, I think it will be helpful for some sample
systems to investigate the impact of EN2 correction in hCI and DhCI calculations (maybe in the
same fashion presented in Figures 3 and 4).
}

\alert{
We agree with the reviewer's suggestion that including the EN2 correction for the excited state calculations is a natural and important addition to the manuscript.
We have performed additional calculations for the full set of states with the different CI models surveyed here (except for the more computationally demanding higher-order models).
These new results were included in Tables II, III, and in the Supporting Information, while a new subsection (IV.E), starting on page x, is devoted to their discussion.
Additions to the Conclusion (page x) and the abstract have also been made.
}

@@

{3) Some of the PES displayed in the SI are not smooth and rather kinky. How does the lack of smoothness affect the fitting procedure and, hence, the vibrational frequencies?
}

\alert{
\ce{CN} is the only system for which some of the potential energy curves are not smooth around the equilibrium geometry.
This is mentioned on page x of the revised manuscript, reading
``For the \ce{CN} radical, CIS, sCI1, and sCI3 produce crossings between ground and excited states around equilibrium, hence non-smooth adiabatic PECs
and somewhat less reliable equilibrium properties in these specific cases.''
}

{4) The authors need to mention (or draw) the molecules of their excited state study. Not everyone might be familiar with the QUEST database. It would be insightful to mention the character of the excited states. I assume these are all singly excited states (no double excitation character?). Can the authors provide additional information regarding their nature (local excitation, charge transfer, Rydberg, etc.)? Specifically, what kind of excited states can be described well or less well with hCI methods?
}

\alert{
The set of systems we considered is now presented in Table II, whreas on page x we describe the character of the excited states, reading
``Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
and 2 doubly-excited states (glyoxal and nitroxyl). It does not include larger molecules nor charge transfer states.''
The details on each specific excited state can be found in The Supporting Information.
New additions to text can be found on page x:
``aa''
The new subsection (IV.E) also discusses how hCI+EN2 models perform for different types of excited states.
}

### 2023-10-22 Manuscript/manuscript.tex
saving work

% Abstract
\begin{abstract}
Hierarchy configuration interaction (hCI) has been recently introduced as an alternative configuration interaction (CI) route combining excitation degree and seniority number, which showed to efficiently recover both dynamic and static correlations for closed-shell molecular systems [\href{https://doi.org/10.1021/acs.jpclett.2c00730}{\textit{J.~Phys.~Chem.~Lett.}~\textbf{2022}, \textit{13}, 4342}].
Here, we generalize hCI for an arbitrary reference determinant, allowing calculations for radicals and for excited states in a state-specific way.
We gauge this route against excitation-based CI (eCI) and seniority-based CI (sCI) by evaluating how different ground-state properties of radicals converge to the full CI limit.
We find that hCI outperforms or matches eCI, whereas sCI is far less accurate, in line with previous observations for closed-shell molecules.
Employing the second-order Epstein-Nesbet perturbation theory as a correction significantly accelerates the convergence of hCI and eCI.
We further explore various hCI and sCI models to calculate excitation energies of closed- and open-shell systems.
Our results underline that both the choice of the reference determinant and the set of orbitals drive the fine balance between correlation of ground and excited states.
State-specific hCI2 and higher order models perform similarly to their eCI counterparts, whereas lower orders of hCI deliver poor results.
In turn, sCI1 produces decent excitation energies for radicals, encouraging the development of related seniority-based coupled cluster methods.
\bigskip
\begin{center}
        \boxed{\includegraphics[keepaspectratio,width=3.25in]{TOC}}
\end{center}
\bigskip
\end{abstract}

@@

The various CI models introduced here were further assessed based on calculated vertical excitation energies for 60 electronic states,
from 18 closed-shell systems and 8 radicals, with geometries extracted from the QUEST database. \cite{Veril_2021}
The full set of excited states and calculated excitation energies, for the various CI models, are provided in the {\SupInf}.
We employed the aug-cc-pVDZ basis set for systems having up to three non-hydrogen atoms and the 6-31+G(d) basis set for the larger ones.
Core orbitals were frozen systematically.
We impose the CI solutions to be eigenstates of the spin angular momentum operator, which implies accounting for a set of appropriate spin-flipped determinants stemming from higher-order excitations or hierarchies. \cite{Chilkuri_2021}
In this sense, our CIS calculations for radicals actually correspond to the so-called extended CIS \cite{Maurice_1996} for instance, and equivalently for the other hCI, eCI, and sCI models.
Notice that spin-contaminated solutions would have different energies than the spin eigenstates considered here.

@@

\begin{comment}
\begin{table}[ht!]
\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE), in Units of eV, with Respect to Reference Theoretical Values, for the Set of
Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
}
\label{tab:stat1}
\begin{ruledtabular}
\begin{tabular}{lddddd}
method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
\hline
CIS               & 48  & +0.03 & 0.61 & 0.59 & 0.77 \\
CISD              & 16  & +4.09 & 4.09 & 4.18 & 0.84 \\
CISDT             & 16  & +0.12 & 0.17 & 0.18 & 0.14 \\
CISDTQ            & 16  & +0.15 & 0.15 & 0.17 & 0.08 \\
%
\hline
hCI1              & 50  & +1.07 & 1.16 & 1.39 & 0.89 \\
hCI1.5            & 50  & +1.95 & 1.95 & 2.04 & 0.59 \\
hCI2              & 16  & +2.99 & 3.53 & 3.61 & 0.76 \\
hCI2.5            & 16  & +1.95 & 1.95 & 2.06 & 0.66 \\
hCI3              & 16  & +0.19 & 0.19 & 0.21 & 0.08 \\
hCI3.5            & 16  & +0.11 & 0.11 & 0.13 & 0.07 \\
%
\hline
$\Delta$CSF       & 50  & -0.71 & 0.77 & 0.91 & 0.58 \\
$\Delta$CISD      & 50  & -0.12 & 0.17 & 0.22 & 0.18 \\
$\Delta$CISDT     & 16  & -0.20 & 0.20 & 0.22 & 0.11 \\
$\Delta$CISDTQ    & 16  & -0.02 & 0.02 & 0.02 & 0.02 \\
%
\hline
$\Delta$hCI1      & 50  & -1.40 & 1.40 & 1.55 & 0.67 \\
$\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15 \\
$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16 \\
$\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 & 0.13 \\
$\Delta$hCI3      & 16  & -0.22 & 0.22 & 0.24 & 0.10 \\
$\Delta$hCI3.5    & 16  & -0.08 & 0.08 & 0.09 & 0.05 \\
%
\hline
sCI2/sCI2         & 50  & +1.35 & 1.35 & 1.51 & 0.68 \\
sCI2/sCI0         & 50  & -0.34 & 0.55 & 0.74 & 0.66 \\
$\Delta$sCI2/sCI2 & 50  & +0.66 & 0.78 & 0.89 & 0.60 \\
$\Delta$sCI2/sCI0 & 50  & -1.04 & 1.04 & 1.21 & 0.62 \\
%
\end{tabular}
\end{ruledtabular}
\end{table}
\end{comment}

\begin{table*}[ht!]
\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE), in Units of eV, with Respect to Reference Theoretical Values, for the Set of
Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
}
\label{tab:stat1}
\begin{ruledtabular}
\begin{tabular}{ldddddlddddd}
method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} & \fk{method}            & \mc{1}{c}{\fk{count}} & \mc{1}{c}{\fk{MSE}} & \mc{1}{c}{\fk{MAE}} & \mc{1}{c}{\fk{RMSE}} & \mc{1}{c}{\fk{SDE}} \\
\hline
CIS               & 48  & +0.03 & 0.61 & 0.59 & 0.77  &\fk{CIS+EN2}               & \fk{48} &       &      &      &      \\
CISD              & 16  & +4.09 & 4.09 & 4.18 & 0.84  & & & & & & \\
CISDT             & 16  & +0.12 & 0.17 & 0.18 & 0.14  & & & & & & \\
CISDTQ            & 16  & +0.15 & 0.15 & 0.17 & 0.08  & & & & & & \\
%
\hline
hCI1              & 50  & +1.07 & 1.16 & 1.39 & 0.89  &\fk{hCI1+EN2}              & \fk{50}  & \fk{+0}.\fk{65}  & \fk{0}.\fk{65} & \fk{0}.\fk{84} & \fk{0}.\fk{53} \\
hCI1.5            & 50  & +1.95 & 1.95 & 2.04 & 0.59  &\fk{hCI1.5+EN2}            & \fk{50}  & \fk{+0}.\fk{62}  & \fk{0}.\fk{63} & \fk{0}.\fk{76} & \fk{0}.\fk{44} \\
hCI2              & 16  & +2.99 & 3.53 & 3.61 & 0.76  & & & & & & \\
hCI2.5            & 16  & +1.95 & 1.95 & 2.06 & 0.66  & & & & & & \\
hCI3              & 16  & +0.19 & 0.19 & 0.21 & 0.08  & & & & & & \\
hCI3.5            & 16  & +0.11 & 0.11 & 0.13 & 0.07  & & & & & & \\
%
\hline
$\Delta$CSF       & 50  & -0.71 & 0.77 & 0.91 & 0.58  &\fk{$\Delta$CSF+EN2}       & \fk{50}  &        &      &      &      \\
$\Delta$CISD      & 50  & -0.12 & 0.17 & 0.22 & 0.18  &\fk{$\Delta$CISD+EN2}      & \fk{50}  & \fk{+0}.\fk{02}  & \fk{0}.\fk{06} & \fk{0}.\fk{09} & \fk{0}.\fk{09} \\
$\Delta$CISDT     & 16  & -0.20 & 0.20 & 0.22 & 0.11  & & & & & & \\
$\Delta$CISDTQ    & 16  & -0.02 & 0.02 & 0.02 & 0.02  & & & & & & \\
%
\hline
$\Delta$hCI1      & 50  & -1.40 & 1.40 & 1.55 & 0.67  &\fk{$\Delta$hCI1+EN2}      & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{24} & \fk{0}.\fk{30} & \fk{0}.\fk{27} \\
$\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15  &\fk{$\Delta$hCI1.5+EN2}    & \fk{50}  & \fk{-0}.\fk{01}  & \fk{0}.\fk{13} & \fk{0}.\fk{19} & \fk{0}.\fk{19} \\
$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16  &\fk{$\Delta$hCI2+EN2}      & \fk{50}  & \fk{+0}.\fk{00}  & \fk{0}.\fk{06} & \fk{0}.\fk{10} & \fk{0}.\fk{09} \\
$\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 & 0.13  & & & & & & \\
$\Delta$hCI3      & 16  & -0.22 & 0.22 & 0.24 & 0.10  & & & & & & \\
$\Delta$hCI3.5    & 16  & -0.08 & 0.08 & 0.09 & 0.05  & & & & & & \\
%
\hline
sCI2/sCI2         & 50  & +1.35 & 1.35 & 1.51 & 0.68  &\fk{sCI2/sCI2+EN2}         & \fk{50}  & \fk{+0}.\fk{60}  & \fk{0}.\fk{60} & \fk{0}.\fk{79} & \fk{0}.\fk{51} \\
sCI2/sCI0         & 50  & -0.34 & 0.55 & 0.74 & 0.66  &\fk{sCI2/sCI0+EN2}         & \fk{50}  & \fk{+0}.\fk{51}  & \fk{0}.\fk{52} & \fk{0}.\fk{69} & \fk{0}.\fk{47} \\
$\Delta$sCI2/sCI2 & 50  & +0.66 & 0.78 & 0.89 & 0.60  &\fk{$\Delta$sCI2/sCI2+EN2} & \fk{50}  & \fk{+0}.\fk{21}  & \fk{0}.\fk{26} & \fk{0}.\fk{35} & \fk{0}.\fk{28} \\
$\Delta$sCI2/sCI0 & 50  & -1.04 & 1.04 & 1.21 & 0.62  &\fk{$\Delta$sCI2/sCI0+EN2} & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{23} & \fk{0}.\fk{27} & \fk{0}.\fk{24} \\
%
\end{tabular}
\end{ruledtabular}
\end{table*}

@@

@@

\begin{comment}
\begin{table}[ht!]
\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE) in Units of eV, with Respect to Reference Theoretical Values, for the Set of
Singly-Excited States from Open-Shell Doublets Listed in the {\SupInf}.
}
\label{tab:stat2}
\begin{ruledtabular}
\begin{tabular}{lddddd}
method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
\hline
CIS               & 19 & +0.38 & 0.41 & 0.63 & 0.50 \\
CISD              &  6 & +2.97 & 2.97 & 3.19 & 1.17 \\
CISDT             &  6 & +0.06 & 0.07 & 0.09 & 0.06 \\
CISDTQ            &  6 & +0.08 & 0.08 & 0.11 & 0.08 \\
%
\hline
hCI1              & 19 & +0.65 & 0.73 & 0.91 & 0.64 \\
hCI1.5            & 19 & +1.07 & 1.07 & 1.26 & 0.68 \\
hCI2              &  6 & +1.32 & 1.32 & 1.60 & 0.90 \\
hCI2.5            &  6 & +0.03 & 0.05 & 0.06 & 0.05 \\
hCI3              &  6 & +0.00 & 0.04 & 0.05 & 0.05 \\
hCI3.5            &  6 & +0.01 & 0.03 & 0.05 & 0.04 \\
%
\hline
$\Delta$CSF       & 19 & -0.04 & 0.43 & 0.59 & 0.59 \\
$\Delta$CISD      & 19 & +0.00 & 0.12 & 0.21 & 0.21 \\
$\Delta$CISDT     &  6 & -0.07 & 0.07 & 0.10 & 0.08 \\
$\Delta$CISDTQ    &  6 & -0.02 & 0.02 & 0.02 & 0.02 \\
%
\hline
$\Delta$hCI1      & 19 & -0.21 & 0.35 & 0.49 & 0.44 \\
$\Delta$hCI1.5    & 19 & -0.14 & 0.45 & 0.70 & 0.69 \\
$\Delta$hCI2      & 19 & -0.02 & 0.10 & 0.16 & 0.16 \\
$\Delta$hCI2.5    &  6 & -0.09 & 0.09 & 0.14 & 0.10 \\
$\Delta$hCI3      &  6 & -0.06 & 0.06 & 0.08 & 0.05 \\
$\Delta$hCI3.5    &  6 & -0.02 & 0.02 & 0.03 & 0.02 \\
%
\hline
sCI1              & 19 & +0.20 & 0.35 & 0.57 & 0.53 \\
$\Delta$sCI1      & 19 & -0.12 & 0.28 & 0.39 & 0.37 \\
\end{tabular}
\end{ruledtabular}
\end{table}
\end{comment}

@@

@@

\fk{
The statistical measures for the CI models corrected with the renormalized EN2 correction are presented in 
Table~\ref{tab:stat1} (closed-shell systems) and Table~\ref{tab:stat2} (open-shell systems),
along with the results for the bare (unperturbed) CI models.
We have not considered the EN2 correction for the higher-order models.
}

\fk{
The EN2 correction decreases the four statistical metrics for both closed- and open-shell systems and across most CI models (the exceptions are discussed in the following).
Importantly, for a given scaling (see Table~\ref{tab:scaling}), whether the bare CI model or the lower-order CI model plus EN2 correction produces the smaller errors depends on the truncation order,
and also on the type of excitations (from closed- or open-shells).
For instance, $\Delta$hCI2 and $\Delta$hCI1+EN2 have the same $\order*{N^6}$ computational cost
but the former has smaller MAEs, specially for the open-shells (\SI{0.10}{\eV} compared to \SI{0.17}{\eV}), but also for the closed-shells (\SI{0.20}{\eV} compared to \SI{0.24}{\eV}).
In turn, $\Delta$hCI1.5+EN2 is significantly more accurate than $\Delta$hCI2.5 for closed-shells, 
with MAEs of \SI{0.13}{\eV} and \SI{0.27}{\eV}, though for open-shells both models produce a MAE of \SI{0.10}{\eV}.
The same is found at the next order.
$\Delta$hCI2+EN2 is much more accurate than $\Delta$hCI3 for closed-shells, with MAEs of \SI{0.06}{\eV} and \SI{0.22}{\eV},
whereas for the open-shell transitions, they deliver comparable MAEs of \SIrange{0.05}{0.06}{\eV}.
%We recall that the small prefactor associated with the calculation of the EN2 correction generally makes it an attractive option.
}

\fk{EN2 with ground-state HF orbitals}

\fk{hCI+EN2 vs eCI+EN2}

\fk{sCI+EN2 vs sCI}

\begin{comment}

\begin{table}[ht!]
\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE), in Units of eV, with Respect to Reference Theoretical Values, for the Set of
Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
}
\label{tab:stat1}
\begin{ruledtabular}
\begin{tabular}{lddddd}
method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} \\
\hline
CIS+EN2               & 48 &       &      &      &      \\
\hline
hCI1+EN2              & 50  & +0.65  & 0.65 & 0.84 & 0.53 \\
hCI1.5+EN2            & 50  & +0.62  & 0.63 & 0.76 & 0.44 \\
%
\hline
$\Delta$CSF+EN2       & 50  &        &      &      &      \\
$\Delta$CISD+EN2      & 50  & +0.02  & 0.06 & 0.09 & 0.09 \\
%
\hline
$\Delta$hCI1+EN2      & 50  & +0.12  & 0.24 & 0.30 & 0.27 \\
$\Delta$hCI1.5+EN2    & 50  & -0.01  & 0.13 & 0.19 & 0.19 \\
$\Delta$hCI2+EN2      & 50  & +0.00  & 0.06 & 0.10 & 0.09 \\
%
\hline
sCI2/sCI2+EN2         & 50  & +0.51  & 0.60 & 0.79 & 0.51 \\
sCI2/sCI0+EN2         & 50  & +0.60  & 0.52 & 0.69 & 0.47 \\
$\Delta$sCI2/sCI2+EN2 & 50  & +0.20  & 0.25 & 0.34 & 0.28 \\
$\Delta$sCI2/sCI0+EN2 & 50  & +0.11  & 0.22 & 0.26 & 0.24 \\
%
\end{tabular}
\end{ruledtabular}
\end{table}

>>>>

% Abstract
\begin{abstract}
Hierarchy configuration interaction (hCI) has been recently introduced as an alternative configuration interaction (CI) route combining excitation degree and seniority number, which showed to efficiently recover both dynamic and static correlations for closed-shell molecular systems [\href{https://doi.org/10.1021/acs.jpclett.2c00730}{\textit{J.~Phys.~Chem.~Lett.}~\textbf{2022}, \textit{13}, 4342}].
Here, we generalize hCI for an arbitrary reference determinant, allowing calculations for radicals and for excited states in a state-specific way.
We gauge this route against excitation-based CI (eCI) and seniority-based CI (sCI) by evaluating how different ground-state properties of radicals converge to the full CI limit.
We find that hCI outperforms or matches eCI, whereas sCI is far less accurate, in line with previous observations for closed-shell molecules.
Employing the second-order Epstein-Nesbet \fk{(EN2)} perturbation theory as a correction significantly accelerates the convergence of hCI and eCI.
We further explore various hCI and sCI models to calculate excitation energies of closed- and open-shell systems.
Our results underline that both the choice of the reference determinant and the set of orbitals drive the fine balance between correlation of ground and excited states.
State-specific hCI2 and higher order models perform similarly to their eCI counterparts, whereas lower orders of hCI deliver poor results,
\fk{unless supplemented by the EN2 correction, which substantially improves their accuracy.}
In turn, sCI1 produces decent excitation energies for radicals, encouraging the development of related seniority-based coupled cluster methods.
\bigskip
\begin{center}
        \boxed{\includegraphics[keepaspectratio,width=3.25in]{TOC}}
\end{center}
\bigskip
\end{abstract}

@@

The various CI models introduced here were further assessed based on calculated vertical excitation energies for 69 electronic states,
from 17 closed-shell systems and 13 radicals \fk{(shown in Table~\ref{tab:systems})}, with geometries extracted from the QUEST database. \cite{Veril_2021}
\fk{Our set includes mostly small systems, ranging from \ce{BH} to glyoxal, displaying a mix of valence and Rydberg singly-excited states,
and 2 doubly-excited states (glyoxal and nitroxyl). It does not include larger molecules nor charge transfer states.}
The full set of excited states and calculated excitation energies, for the various CI models, are provided in the {\SupInf}.

\begin{table}[ht!]
\caption{\fk{Systems Considered in the Present Excited State Calculations.}}
\label{tab:systems}
{\color{orange}
\begin{ruledtabular}
\begin{tabular}{lll|lll}
\mc{3}{c|}{Closed-shells}           & \mc{3}{c}{Open-shells} \\
\hline
glyoxal      & \ce{HCF}  & \ce{H2O} & \ce{BeH} & \ce{NH2} & \ce{HCO} \\
acetaldehyde & \ce{HCCl} & \ce{H2S} & \ce{BeF} & \ce{PH2} & \ce{HOC} \\
silylidene   & \ce{HPO}  & \ce{N2}  & \ce{BH2} & vinyl    & \ce{OH}  \\
nitroxyl     & \ce{CF2}  & \ce{NH3} & \ce{CN}  & allyl    & \ce{CO+} \\
ethylene     & \ce{BH}   & \ce{HCl} & \ce{CH3} &          & \\
methanimine  & \ce{BF}   &          &          &          & \\
\end{tabular}
\end{ruledtabular}
}
\end{table}

We employed the aug-cc-pVDZ basis set for systems having up to three non-hydrogen atoms and the 6-31+G(d) basis set for the larger ones.
Core orbitals were frozen systematically.
We impose the CI solutions to be eigenstates of the spin angular momentum operator, which implies accounting for a set of appropriate spin-flipped determinants stemming from higher-order excitations or hierarchies. \cite{Chilkuri_2021}
In this sense, our CIS calculations for radicals actually correspond to the so-called extended CIS \cite{Maurice_1996} for instance, and equivalently for the other hCI, eCI, and sCI models.
Notice that spin-contaminated solutions would have different energies than the spin eigenstates considered here.

@@

\begin{table*}[ht!]
\caption{Mean Signed Error (MSE), Mean Absolute Error (MAE), Root-Mean Square Error (RMSE), and Standard Deviation of the Errors (SDE), in Units of eV, with Respect to Reference Theoretical Values, for the Set of
Singly Excited States of Closed-Shell Systems Listed in the {\SupInf}.
}
\label{tab:stat1}
\begin{ruledtabular}
\begin{tabular}{ldddddlddddd}
method            & \mc{1}{c}{count} & \mc{1}{c}{MSE} & \mc{1}{c}{MAE} & \mc{1}{c}{RMSE} & \mc{1}{c}{SDE} & \fk{method}            & \mc{1}{c}{\fk{count}} & \mc{1}{c}{\fk{MSE}} & \mc{1}{c}{\fk{MAE}} & \mc{1}{c}{\fk{RMSE}} & \mc{1}{c}{\fk{SDE}} \\
\hline
CIS               & 48  & +0.03 & 0.61 & 0.59 & 0.77  &\fk{CIS+EN2}               & \fk{48} &       &      &      &      \\
CISD              & 16  & +4.09 & 4.09 & 4.18 & 0.84  & & & & & & \\
CISDT             & 16  & +0.12 & 0.17 & 0.18 & 0.14  & & & & & & \\
CISDTQ            & 16  & +0.15 & 0.15 & 0.17 & 0.08  & & & & & & \\
%
\hline
hCI1              & 50  & +1.07 & 1.16 & 1.39 & 0.89  &\fk{hCI1+EN2}              & \fk{50}  & \fk{+0}.\fk{65}  & \fk{0}.\fk{65} & \fk{0}.\fk{84} & \fk{0}.\fk{53} \\
hCI1.5            & 50  & +1.95 & 1.95 & 2.04 & 0.59  &\fk{hCI1.5+EN2}            & \fk{50}  & \fk{+0}.\fk{62}  & \fk{0}.\fk{63} & \fk{0}.\fk{76} & \fk{0}.\fk{44} \\
hCI2              & 16  & +2.99 & 3.53 & 3.61 & 0.76  & & & & & & \\
hCI2.5            & 16  & +1.95 & 1.95 & 2.06 & 0.66  & & & & & & \\
hCI3              & 16  & +0.19 & 0.19 & 0.21 & 0.08  & & & & & & \\
hCI3.5            & 16  & +0.11 & 0.11 & 0.13 & 0.07  & & & & & & \\
%
\hline
$\Delta$CSF       & 50  & -0.71 & 0.77 & 0.91 & 0.58  &\fk{$\Delta$CSF+EN2}       & \fk{50}  &        &      &      &      \\
$\Delta$CISD      & 50  & -0.12 & 0.17 & 0.22 & 0.18  &\fk{$\Delta$CISD+EN2}      & \fk{50}  & \fk{+0}.\fk{02}  & \fk{0}.\fk{06} & \fk{0}.\fk{09} & \fk{0}.\fk{09} \\
$\Delta$CISDT     & 16  & -0.20 & 0.20 & 0.22 & 0.11  & & & & & & \\
$\Delta$CISDTQ    & 16  & -0.02 & 0.02 & 0.02 & 0.02  & & & & & & \\
%
\hline
$\Delta$hCI1      & 50  & -1.40 & 1.40 & 1.55 & 0.67  &\fk{$\Delta$hCI1+EN2}      & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{24} & \fk{0}.\fk{30} & \fk{0}.\fk{27} \\
$\Delta$hCI1.5    & 50  & -2.80 & 2.80 & 3.03 & 1.15  &\fk{$\Delta$hCI1.5+EN2}    & \fk{50}  & \fk{-0}.\fk{01}  & \fk{0}.\fk{13} & \fk{0}.\fk{19} & \fk{0}.\fk{19} \\
$\Delta$hCI2      & 50  & -0.18 & 0.20 & 0.25 & 0.16  &\fk{$\Delta$hCI2+EN2}      & \fk{50}  & \fk{+0}.\fk{01}  & \fk{0}.\fk{07} & \fk{0}.\fk{10} & \fk{0}.\fk{10} \\
$\Delta$hCI2.5    & 16  & -0.27 & 0.27 & 0.30 & 0.13  & & & & & & \\
$\Delta$hCI3      & 16  & -0.22 & 0.22 & 0.24 & 0.10  & & & & & & \\
$\Delta$hCI3.5    & 16  & -0.08 & 0.08 & 0.09 & 0.05  & & & & & & \\
%
\hline
sCI2/sCI2         & 50  & +1.35 & 1.35 & 1.51 & 0.68  &\fk{sCI2/sCI2+EN2}         & \fk{50}  & \fk{+0}.\fk{60}  & \fk{0}.\fk{60} & \fk{0}.\fk{79} & \fk{0}.\fk{51} \\
sCI2/sCI0         & 50  & -0.34 & 0.55 & 0.74 & 0.66  &\fk{sCI2/sCI0+EN2}         & \fk{50}  & \fk{+0}.\fk{51}  & \fk{0}.\fk{52} & \fk{0}.\fk{69} & \fk{0}.\fk{47} \\
$\Delta$sCI2/sCI2 & 50  & +0.66 & 0.78 & 0.89 & 0.60  &\fk{$\Delta$sCI2/sCI2+EN2} & \fk{50}  & \fk{+0}.\fk{21}  & \fk{0}.\fk{26} & \fk{0}.\fk{35} & \fk{0}.\fk{28} \\
$\Delta$sCI2/sCI0 & 50  & -1.04 & 1.04 & 1.21 & 0.62  &\fk{$\Delta$sCI2/sCI0+EN2} & \fk{50}  & \fk{+0}.\fk{12}  & \fk{0}.\fk{23} & \fk{0}.\fk{27} & \fk{0}.\fk{24} \\
%
\end{tabular}
\end{ruledtabular}
\end{table*}

@@

\fk{[Performance of the hCI models for the different types of transitions.]}

@@

@@

\fk{[Performance of the sCI models for the different types of transitions.]}

@@

\fk{
The renormalized EN2 correction significantly reduces the errors across most CI models, for both closed- and open-shell systems.
The statistical measures for the CI models corrected with the EN2 correction can be seen in
Table~\ref{tab:stat1} (closed-shell systems) and Table~\ref{tab:stat2} (open-shell systems),
along with the results for the bare (unperturbed) CI models.
We have not considered the EN2 correction for the higher-order models.
}

\fk{
For a given scaling (see Table~\ref{tab:scaling}), whether the bare CI model or the lower-order CI model plus EN2 correction produces the smaller errors depends on the truncation order,
and also on the type of excitations (from closed- or open-shells).
For instance, $\Delta$hCI2 and $\Delta$hCI1+EN2 have the same $\order*{N^6}$ computational cost
but the former has smaller MAEs, for closed-shells (\SI{0.20}{\eV} compared to \SI{0.24}{\eV}), and more significantly for open-shells (\SI{0.10}{\eV} compared to \SI{0.17}{\eV}).
In turn, $\Delta$hCI1.5+EN2 is significantly more accurate than $\Delta$hCI2.5 for closed-shells,
with MAEs of \SI{0.13}{\eV} and \SI{0.27}{\eV}, whreas for open-shells both models produce a MAE of \SI{0.10}{\eV}.
The same is found at the next order.
$\Delta$hCI2+EN2 is much more accurate than $\Delta$hCI3 for closed-shells, with MAEs of \SI{0.06}{\eV} and \SI{0.22}{\eV},
whereas for the open-shell transitions, they deliver comparable MAEs of \SIrange{0.05}{0.06}{\eV}.
We recall that the calculation of the EN2 correction has a small associated prefactor, which makes the above $\Delta$hCI+EN2 models attractive alternatives to their higher-order unperturbed $\Delta$hCI counterparts.
}

\fk{
Among the different CI models, it is worth highlighting the huge improvement observed from $\Delta$hCI1.5 to $\Delta$hCI1.5+EN2.
For the closed-shells, the MAE drops from \SI{2.80}{\eV} to \SI{0.13}{\eV}, whereas a reduction from \SI{0.45}{\eV} to \SI{0.10}{\eV} is seen for the open-shells.
In both cases, the MSE is virtually zero.
In addition, the individual errors for singlets, triplets, Rydberg, and valence states are comparable, lying in the range \SIrange{0.11}{0.15}{\eV}.
}

\fk{
We found a similar impact of the EN2 correction for $\Delta$CISD and $\Delta$hCI2.
As discussed above, $\Delta$CISD is somewhat more accurate than $\Delta$hCI2 for the closed-shells (MAEs of \SI{0.17}{\eV} and \SI{0.20}{\eV}),
whereas the EN2 correction decreases their respective MAE to \SI{0.06}{\eV} and \SI{0.07}{\eV}.
In turn, while $\Delta$CISD is slightly less accurate than $\Delta$hCI2 for the open-shells (MAEs of \SI{0.12}{\eV} and \SI{0.10}{\eV}),
the EN2 corrected models show a MAE of \SI{0.04}{\eV} and \SI{0.05}{\eV}, respectively.
Furthermore, the accuracy of $\Delta$hCI2+EN2 is comparable for singlets and triplets (MAE of \SIrange{0.06}{0.07}{\eV}),
and somewhat superior for Rydberg states when compared to valence states (MAE of \SI{0.04}{\eV} and \SI{0.08}{\eV}).
The same trend is found for $\Delta$CISD+EN2, which slightly outperforms $\Delta$hCI2+EN2 for each of the four types of transitions mentioned above.
}

\fk{
In another contrast to eCI and hCI routes, the EN2 correction has a more limited effect to the sCI route.
There is little to no advance for the models based on ground-state HF orbitals.
The outcome is more favourable with state-specific orbitals,
where, for the closed-shells, $\Delta$sCI2/sCI2+EN2 and $\Delta$sCI2/sCI0+EN2 show comparable MAEs (\SI{0.26}{\eV} and \SI{0.23}{\eV}, respectively).
For the open-shells, the MAE decrease from \SI{0.28}{\eV} ($\Delta$sCI1) to \SI{0.12}{\eV} ($\Delta$sCI1+EN2),
the latter error being comparable to that of the much less expensive $\Delta$hCI1.5+EN2 model (\SI{0.10}{\eV}).
Despite the improvement, the EN2 correction is not enough to render the sCI models competitive.
}
